---
title: "DA R PROGRAMMING EXAM"
author: "DA class"
date: "22/7/2024"
output: html_document
---

 ## IMPORT THE DATASET CALLED employees_attrition.csv file 
 ## INTO RSTUDIO
 
 ## EMPLOYEE ATTRITION MEANS EMPLOYEES ARE LEAVING THE COMPANY
 ## FOR ONE REASON OR THE OTHER.
 
 ## THE ATTRITION COLUMN HAS TWO POSSIBLE VALUES YES - FOR EMPLOYEES
 ## THAT LEFT THE COMPANY AND NO FOR THOSE THAT ARE  STILL AROUND.
 
 ## NOTE
 
 ## 1. RUN EVERY CODE CHUNCK YOU CREATE SO THAT THE OUTPUT WILL
 ## BE DISPLAYED BELOW THE CODE CHUNK.
 
 ## 2. SAVE YOUR RMARKDOWN FILE WITH YOUR NAME FOR IDENTIFICATION
 
 ## 3. USE DPLYR FUNCTIONS TO ANSWER THE FOLLOWING QUESTIONS.
 
 
 
```{r Q1}

# load the tidyverse package into this session
library("tidyverse")
```
```{r}
library("tidyverse")
```
 
```{r  Q2}
## import the given employees_attrition.csv file 

```
```{r}
employees_attestation <- read.csv(file.choose())
```

```{r Q3 }

## SHOW THE STRUCTURE OF THE DATAFRAME

```

```{r}
str(employees_attestation)
```

```{r Q4 }

## HOW MANY ROWS AND COLUMNS ARE THERE IN THE DATAFRAME

```
```{r}
view(employees_attestation)
```
```{r}
dim(employees_attestation)
``
``##`{r Q5}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY AND THE 
## NUMBER THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function to categorise.
```

```

```
```{r}
employees_attestation%>%
  filter(attrition %in% c("Yes","No"))
group_by(employees_attestation,attrition)%>%
  summarise(count=n())
```


```{r Q6}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY IN EACH 
## DEPARTMENT.

```
```{r}
employees_attestation%>%
  filter(attrition=="Yes")%>%
group_by(attrition,department)%>%
  summarise(count=n())
```



```{r Q7}
## SHOW THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT LEFT THE COMPANY
## AND THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function



```
```{r}
employees_attestation%>%
 group_by(gender,attrition)%>%
  summarise(count=n())
```




```{r Q8}

## WHAT IS THE AVERAGE NUMBER OF YEARS WORKED IN THE COMPANY BY THOSE THAT LEFT THE COMPANY.

```
```{r}
employees_attestation%>%
  filter(attrition=="Yes")%>%
  group_by(attrition)%>%
  summarise(avr_year_worked=mean(YearsAtCompany))
  
```



```{r Q9}
##using GGPLOT AND  DPLYR, VISUALISE USING COLUMNS CHART ,THE 5 JOB ROLES WITH THE HIGEHST NUMBER OF EMPLOYESS THAT LEFT .
##HINT USE ggplot(aes(x =val,y=val,fill=val))+
##geom_col()
```

```{r}
employees_attestation%>%
  group_by(job_role)%>%
  summarise(count=n())%>%
arrange(desc(count))%>%
  head(5)%>%
  ggplot(aes(x=reorder(job_role,count) ,y =count ,fill =job_role)) +
geom_col() +
coord_flip()
```



```{r Q10}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY BASED ON  THE isMarried column.
```
```{r}
movies
```
```{r}
employees_attestation%>%
  filter(attrition=="Yes")%>%
  group_by(isMarried)%>%
    summarise(count=n())

```


```{r Q10}
```



```{r Q11 }
## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO LEFT 
## THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS.



```
```{r}
employees_attestation%>%
  filter(attrition=="Yes")%>%
  group_by(department)%>%
  summarise(avg_incom = mean(monthly_income))%>%
  arrange(-avg_incom)
```


```{r Q12}

## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO ARE STILL WITH THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS.

```
```{r}
employees_attestation%>%
  filter(attrition=="No")%>%
  group_by(department)%>%
  summarise(avg_incom = mean(monthly_income))%>%
  arrange(-avg_incom)
```



